\LaTeX\ Author Guidelines for ICCV Proceedings

### 2022-05-19 chapters/0_abstract.tex
Integrate last Fred's comments
Camera ready changes

% !TEX root = ../agglo_clust_review.tex
%%%%%%%%% ABSTRACT
\begin{abstract}
We propose a simple theoretical framework that generalizes simple and fast algorithms for hierarchical agglomerative clustering to weighted graphs with both attractive and repulsive interactions between the nodes. This framework defines GASP, a Generalized Algorithm for Signed graph Partitioning, and allows us to explore many combinations of different linkage criteria and cannot-link constraints. 
We prove the equivalence of existing clustering methods to some of those combinations and introduce new algorithms for combinations that have not been studied before. 
We study both theoretical and empirical properties of these combinations and prove which of these define an ultrametric on the graph.
We conduct a systematic comparison of various instantiations of GASP on a large variety of both synthetic and existing signed clustering problems, in terms of accuracy but also efficiency and robustness to noise. 
Lastly, we find that some of the algorithms included in our framework, when combined with the predictions from a CNN model, result in a simple bottom-up instance segmentation pipeline.
Going all the way from pixels to final segments with a simple procedure, we achieve state-of-the-art accuracy on the CREMI 2016 EM segmentation benchmark without requiring domain-specific superpixels.
% On stochastic block model problems, GASP compares favorably to spectral clustering. 
% \keywords{Partitioning of Signed Graphs, Generalized Framework, Agglomerative Clustering, Instance Segmentation}
\end{abstract}

>>>>

% !TEX root = ../agglo_clust_review.tex
%%%%%%%%% ABSTRACT
\begin{abstract}
We propose a theoretical framework that generalizes simple and fast algorithms for hierarchical agglomerative clustering to weighted graphs with both attractive and repulsive interactions between the nodes. This framework defines GASP, a Generalized Algorithm for Signed graph Partitioning\footnote{Code available at: \url{https://github.com/abailoni/GASP}}, and allows us to explore many combinations of different linkage criteria and cannot-link constraints. 
We prove the equivalence of existing clustering methods to some of those combinations and introduce new algorithms for combinations that have not been studied before. 
We study both theoretical and empirical properties of these combinations and prove that some of these define an ultrametric on the graph.
We conduct a systematic comparison of various instantiations of GASP on a large variety of both synthetic and existing signed clustering problems, in terms of accuracy but also efficiency and robustness to noise. 
Lastly, we show that some of the algorithms included in our framework, when combined with the predictions from a CNN model, result in a simple bottom-up instance segmentation pipeline.
Going all the way from pixels to final segments with a simple procedure, we achieve state-of-the-art accuracy on the CREMI 2016 EM segmentation benchmark without requiring domain-specific superpixels.
% On stochastic block model problems, GASP compares favorably to spectral clustering. 
% \keywords{Partitioning of Signed Graphs, Generalized Framework, Agglomerative Clustering, Instance Segmentation}
\end{abstract}

### 2022-05-19 chapters/3_algorithm.tex
Camera ready changes

\textbf{Linkage criteria and hierarchical trees} -- 
% We also denote as $S_u$ the cluster associated with node $u$ \TODO{Needed?}.
% We call two clusters $S,S'$ \emph{adjacent} if there exists at least one edge ${e_{ts}\in E}$ connecting a node $t\in S_u$ to a node $s\in S_v$. 
Let the interaction $\interact(S,S')\in\mathbb{R}$ between two clusters $S,S'$ be defined as a function, named \emph{linkage criterion}, depending on the weights of \emph{all} edges connecting clusters $S$ and $S'$, i.e. ${(S \times S') \cap E}$ where $\times$ denotes the outer product. 
The linkage criteria tested in this article are listed and defined in Table \ref{tab:linkage-criteria}.
A \emph{dendrogram} $T$ is a rooted binary tree\footnote{In general, one could look at trees that are not binary. However, the algorithms discussed in this paper always generate binary hierarchical trees, so nothing would be gained by this generalization.} representing the merging order of an agglomerative algorithm, such that the leaves of the tree are in one-to-one correspondence with $V$ and each node of the tree represents a merge between two clusters. 
Let $T_{\mathrm{R}},T_{\mathrm{L}}\subset T$ denote the subtrees rooted at the two children of the root node in $T$.
For any two leaves $u,v \in V$, let $T[u \vee v]$ be the subtree rooted at the least common ancestor $(u \vee v)\in T$ of nodes $u$ and $v$ (furthest from the root), and let \texttt{leaves}$(T[u \vee v])\subseteq V$ be the set of leaves of this subtree. 
% \TODO{Here we talk about the algorithm already, so we could move it to next section and the flow could be better. But we need a tree in the alg. So we better do this later}
% Algorithm \ref{main_alg} returns a binary tree denoted by $T^*$.
Given an agglomerative algorithm with merging tree $T$, let $h_T:V \times V \rightarrow \mathbb{N}$ denote the \emph{dendrogram-height} of each $(u\vee v)\in T$, which is defined as the iteration number at which nodes $u,v\in V$ were merged by the algorithm (see example in Fig.~\ref{fig:intro_figure}a). We also define $\treeHeight(u,v)$ as the signed interaction $\interact{}(S,S')$ between the two clusters $S,S'$ that were merged at iteration $h_T(u, v)$: 
% Then, given a binary tree $T$ and a linkage criterion $\interact$, we can associate a \emph{signed similarity} function $\treeHeight$ to each pair of distinct original nodes $u\neq v$, with $u, v \in V$:
\begin{equation}\label{eq:def_dendr_interact}
\treeHeight(u,v) \equiv \interact{} \big( \text{\texttt{leaves}}(T_{\mathrm{R}}[u \vee v]), \text{\texttt{leaves}}(T_{\mathrm{L}}[u \vee v]) \big)
\end{equation}
% \OPTIONAL{In other words, $h_T(u,v)$ and $\treeHeight(u,v)$ represent the \emph{dendrogram-height} and \emph{signed-interaction} associated to each $u \vee v$ node in the tree $T$ (see toy example in \TODO{Fig.~1}).}

@@

In phase one, \algname{} selects the pair of clusters with the highest absolute interaction $|\interact(S, S')|$, so that the most attractive and the most repulsive pairs are analyzed first. If the interaction is repulsive and the algorithm option \emph{addCannotLinkConstraints} is \texttt{True}, then the two clusters are constrained so that their members can never merge in subsequent steps of phase one. If the interaction is attractive, then the clusters are merged, provided that they were not previously constrained. 
After each merge, the interaction between the merged cluster and its neighbors is updated according to one of the linkage criteria $\interact(S, S')$ listed in Table \ref{tab:linkage-criteria}. Phase one terminates when all the remaining clusters are either constrained or share repulsive interactions. Note that, on unsigned graphs, in phase one all nodes are merged into a single cluster and \algname{} is then equivalent to a standard hierarchical clustering algorithm.

@@

\begin{algorithm}[t]
\footnotesize
  \begin{flushleft}
  \footnotesize
  \caption{\algname{}}
  % \caption{\algname{}: generalized algorithm for signed graph partitioning}
   \hspace*{\algorithmicindent} \textbf{Input:} Graph $\mathcal{G}(V,E,w^+,w^-)$; linkage criterion $\interact{}$; \\ 
   \hspace*{4.3em}boolean {\color{blue}addCannotLinkConstraints}  \\
  \hspace*{\algorithmicindent} \textbf{Output:} Final clustering $\Pi^*$, rooted binary hierarchical tree $T^*$\\
  \hspace*{\algorithmicindent} 
  \begin{algorithmic}[1]
  \footnotesize
  % \small
      \State Initial clustering: $\Pi=\{\{v_1\}, \ldots, \{v_{|V|}\}\}$
      \State Initialize hierarchical tree $T^*$ with leaf nodes  $V=\{v_1,\ldots,v_{|V|}\}$
      \State Initialize  cluster interactions with $\cost_e = w^+_e - w^-_e$, $\forall e\in E$
      \State \emph{// Phase 1: Merge positive interactions (possibly using constraints)}
      \State Push incident nodes of every edge $e\in E$ to priority queue (PQ) with priority $|w_e|$
      \Repeat 
        \State Pop $S,S'\in\Pi$ with highest interaction $|\interact{}(S,S')|$ from PQ
        \If{\big[{\color{green}\textbf{$\interact{}(S,S') > 0$}}\big] \textbf{and} \big[$S,S'$ \textbf{not} constrained\big]}
          \State Merge clusters $S$, $S'$ and update hierarchical tree $T^*$
          \State Update interactions \& constraints with neighboring clusters
        \ElsIf{{\color{blue}addCannotLinkConstr} \textbf{and}  \big[{\color{red}\textbf{$\interact{}(S,S') \leq 0$}}\big]}
          \State Add CannotLink Constraint between $S$ and $S'$
        \EndIf
      \Until{\big[$PQ$ is empty\big]}
      % \State
      % \If{{\color{blue}addCannotLinkConstr}} % \Comment{Step 2: release constraints}
      \State \emph{// Phase 2: Remove constraints \& merge all positive interactions}
      \State Push signed interactions $\interact{}(S,S')$ to PQ, $\forall S, S' \in \Pi$
      \Repeat 
        \State Pop $S,S'\in\Pi$ with highest interaction $\interact{}(S,S')$ from PQ
        \If{\big[{\color{green}\textbf{$\interact{}(S,S') > 0$}}\big]}
          \State Merge clusters $S$, $S'$ and update hierarchical tree $T^*$
          \State Update interactions with neighboring clusters
        \EndIf
      \Until{\big[{\color{red}\textbf{$\interact{}(S,S') \leq 0$}}\big]} 
      \State Save the final clustering $\Pi^* \gets \Pi$ 
      \State \emph{// Phase 3: Merge negative interactions until one single cluster is left}
      \Repeat
        \State Pop $S,S'\in\Pi$ with highest interaction $\interact{}(S,S')$ from PQ 
        \State Merge clusters $S$, $S'$ and update hierarchical tree $T^*$
        \State Update interactions with neighboring clusters
      \Until{\big[Only one cluster is left in $\Pi$\big]} 
      % \EndIf
      \State
      \Return $\Pi^*$, $T^*$
  \end{algorithmic}
    \label{main_alg}
  \end{flushleft}

@@

\paragraph{Average, Single, and Complete Linkage} -- These three linkage criteria have been thoroughly studied on unsigned graphs, but never - until very recently - on signed graphs. In concurrent independent related work \cite{chehreghani2020hierarchical}, the authors prove that applying these three linkage methods to a signed graph is equivalent to applying them to the unsigned graph obtained by shifting all edge weights by a constant.  Here, we prove which of the algorithms studied here are ``intrinsically signed'' and do not have this invariance-property:
\begin{restatable}{prop}{invariantAlgs}
\label{prop:weight_shift_invariant}
We call an agglomerative algorithm ``weight-shift invariant'' if the dendrogram $T$ returned by the algorithm is invariant w.r.t. a shift of all edge weights $w_e$  by a constant $\alpha\in \mathbb{R}$. Among the variations of \algname{}, only hierarchical clustering with Average (HC-Avg), Single (HC-Single), and Complete linkage (HC-Complete) are weight-shift-invariant (see green box in Table~\ref{tab:linkage-criteria}).
\end{restatable}
\noindent Although average and single linkage methods have been largely studied on unsigned graphs, to our knowledge, they have never been combined with cannot-link constraints on signed graphs\footnote{Note that Complete linkage methods return the same clustering whether constraints are enforced or not (proof in Lemma \ref{lemma:absMax_and_complete_property}, in Appendix).}, so we name these algorithms \emph{HCC-Avg} and \emph{HCC-Single}. 
% On the other hand, we prove that an agglomerative algorithm with Complete linkage (HC-Complete) returns the same clustering whether cannot-link-constraints are enforced or not (Lemma \ref{lemma:absMax_and_complete_property}, in Appendix).
% , we prove that with Complete linkage the agglomeration order does not change whether constraints are introduced or not.

@@

\paragraph{Algorithms defining an ultrametric} -- The connection between agglomerative algorithms and ultrametrics\footnote{A metric space $(X,d)$ is an \emph{ultrametric} if, for every $x,y,z \in X$, $d(x,y)\leq \max \{d(x,z), d(y,z)\}$.} is well known. Usually, ultrametrics are associated to strictly positive \emph{similarity} or \emph{dissimilarity} measures on a graph. In our framework, a trivial ultrametric is always given by the height $h_T$ of the dendrogram. However, for some of the \algname{} variations, we now define an ultrametric based on the edge weights and the signed interactions between clusters, generalizing what has been done for HC on unsigned graphs \cite{johnson1967hierarchical,milligan1979ultrametric}. To define this measure and prove its ultrametric property, we first map the signed interaction $\treeHeight{}$ defined in Eq.~\ref{eq:def_dendr_interact} to positive ``pseudo-distances'' $d_{T}:V \times V \rightarrow \mathbb{R}^{+}$:
\begin{align}\label{eq:UM_def}
d_{T}(u,v)\equiv &\begin{cases}
0 & \text{if}\,\, u=v\\
M-\treeHeight(u,v) & \text{if}\,\, u\neq v
\end{cases}\quad \forall u,v\in V\\
\mathrm{where}& \quad  M \equiv  \epsilon + \max_{u',v'\in V,\,u'\neq v'}\treeHeight(u',v')
\end{align}
and where $\epsilon > 0$. 
% By definition, we have that $d_{T}(u,v)=d_{T}(v,u)$ and $d_{T}(u,u)= 0$. 
We then prove the following proposition:
\begin{restatable}{prop}{secondUltraMetricProperty}
\label{prop:ultraMetric2}
% Consider the binary rooted tree $T^*$ returned by the \TODO{all variants in Table 1} \algname{} Algorithm~\ref{main_alg}, given a choice of linkage criteria $\interact{}$ and the use (or not) of CannotLinkConstraints. 
Among the algorithms included in the \algname{} framework (see Table \ref{tab:linkage-criteria}), only Mutex Watershed and hierarchical clustering with Average (HC-Avg), Single (HC-Single) and Complete linkage (HC-Complete) define an ultrametric $(V, d_{T^*})$, where $d_{T^*}$ is defined in Eq.~\ref{eq:UM_def} and $T^*$ is the tree returned by the \algname{} Algorithm~\ref{main_alg}.
\end{restatable}
\noindent In summary, in this section we have extended the family of HC algorithms \cite{johnson1967hierarchical,milligan1979ultrametric} with ``weight-based ultrametrics'' to signed graphs. Next, we move to their empirical evaluation.

>>>>

\textbf{Linkage criteria and hierarchical trees} -- 
% We also denote as $S_u$ the cluster associated with node $u$ \TODO{Needed?}.
% We call two clusters $S,S'$ \emph{adjacent} if there exists at least one edge ${e_{ts}\in E}$ connecting a node $t\in S_u$ to a node $s\in S_v$. 
Let the interaction $\interact(S \cup S')\in\mathbb{R}$ between two clusters $S,S'$ be defined as a function, named \emph{linkage criterion}, depending on the weights of \emph{all} edges connecting clusters $S$ and $S'$.
% , i.e. ${(S \times S') \cap E}$ where $\times$ denotes the outer product. 
The linkage criteria tested in this article are listed and defined in Table \ref{tab:linkage-criteria}.
A \emph{dendrogram} $T$ is a rooted binary tree\footnote{In general, one could look at trees that are not binary. However, the algorithms discussed in this paper always generate binary hierarchical trees, so nothing would be gained by this generalization.} representing the merging order of an agglomerative algorithm, such that the leaves of the tree are in one-to-one correspondence with $V$ and each node of the tree represents a merge between two clusters. 
Let $T_{\mathrm{R}},T_{\mathrm{L}}\subset T$ denote the subtrees rooted at the two children of the root node in $T$.
For any two leaves $u,v \in V$, let $T[u \vee v]$ be the subtree rooted at the least common ancestor $(u \vee v)\in T$ of nodes $u$ and $v$ (furthest from the root), and let \texttt{leaves}$(T[u \vee v])\subseteq V$ be the set of leaves of this subtree. 
% \TODO{Here we talk about the algorithm already, so we could move it to next section and the flow could be better. But we need a tree in the alg. So we better do this later}
% Algorithm \ref{main_alg} returns a binary tree denoted by $T^*$.
Given an agglomerative algorithm with merging tree $T$, let $h_T:V \times V \rightarrow \mathbb{N}$ denote the \emph{dendrogram-height} of each $(u\vee v)\in T$, which is defined as the iteration number at which nodes $u,v\in V$ were merged by the algorithm (see example in Fig.~\ref{fig:intro_figure}a). We also define $\treeHeight(u,v)$ as the signed interaction $\interact{}(S \cup S')$ between the two clusters $S,S'$ that were merged at iteration $h_T(u, v)$: 
% Then, given a binary tree $T$ and a linkage criterion $\interact$, we can associate a \emph{signed similarity} function $\treeHeight$ to each pair of distinct original nodes $u\neq v$, with $u, v \in V$:
\begin{equation}\label{eq:def_dendr_interact}
\treeHeight(u,v) \equiv \interact{} \big( \text{\texttt{leaves}}(T_{\mathrm{R}}[u \vee v])  \cup  \text{\texttt{leaves}}(T_{\mathrm{L}}[u \vee v]) \big)
\end{equation}
% \OPTIONAL{In other words, $h_T(u,v)$ and $\treeHeight(u,v)$ represent the \emph{dendrogram-height} and \emph{signed-interaction} associated to each $u \vee v$ node in the tree $T$ (see toy example in \TODO{Fig.~1}).}

@@

In phase one, \algname{} selects the pair of clusters with the highest absolute interaction $|\interact(S  \cup  S')|$, so that the most attractive and the most repulsive pairs are analyzed first. If the interaction is repulsive and the algorithm option \emph{addCannotLinkConstraints} is \texttt{True}, then the two clusters are constrained so that their members can never merge in subsequent steps of phase one. If the interaction is attractive, then the clusters are merged, provided that they were not previously constrained. 
After each merge, the interaction between the merged cluster and its neighbors is updated according to one of the linkage criteria $\interact(S \cup  S')$ listed in Table \ref{tab:linkage-criteria}. Phase one terminates when all the remaining clusters are either constrained or share repulsive interactions. Note that, on unsigned graphs, in phase one all nodes are merged into a single cluster and \algname{} is then equivalent to a standard hierarchical clustering algorithm.

@@

\begin{algorithm}[t]
\footnotesize
  \begin{flushleft}
  \footnotesize
  \caption{\algname{}}
  % \caption{\algname{}: generalized algorithm for signed graph partitioning}
   \hspace*{\algorithmicindent} \textbf{Input:} Graph $\mathcal{G}(V,E,w^+,w^-)$; linkage criterion $\interact{}$; \\ 
   \hspace*{4.3em}boolean {\color{blue}addCannotLinkConstraints}  \\
  \hspace*{\algorithmicindent} \textbf{Output:} Final clustering $\Pi^*$, rooted binary hierarchical tree $T^*$\\
  \hspace*{\algorithmicindent} 
  \begin{algorithmic}[1]
  \footnotesize
  % \small
      \State Initial clustering: $\Pi=\{\{v_1\}, \ldots, \{v_{|V|}\}\}$
      \State Initialize hierarchical tree $T^*$ with leaf nodes  $V=\{v_1,\ldots,v_{|V|}\}$
      \State Initialize  cluster interactions with $\cost_e = w^+_e - w^-_e$, $\forall e\in E$
      \State \emph{// Phase 1: Merge positive interactions (possibly using constraints)}
      \State Push incident nodes of every edge $e\in E$ to priority queue (PQ) with priority $|w_e|$
      \Repeat 
        \State Pop $S,S'\in\Pi$ with highest interaction $|\interact{}(S  \cup  S')|$ from PQ
        \If{\big[{\color{green}\textbf{$\interact{}(S \cup S') > 0$}}\big] \textbf{and} \big[$S,S'$ \textbf{not} constrained\big]}
          \State Merge clusters $S$, $S'$ and update hierarchical tree $T^*$
          \State Update interactions \& constraints with neighboring clusters
        \ElsIf{{\color{blue}addCannotLinkConstr} \textbf{and}  \big[{\color{red}\textbf{$\interact{}(S \cup  S') \leq 0$}}\big]}
          \State Add CannotLink Constraint between $S$ and $S'$
        \EndIf
      \Until{\big[$PQ$ is empty\big]}
      % \State
      % \If{{\color{blue}addCannotLinkConstr}} % \Comment{Step 2: release constraints}
      \State \emph{// Phase 2: Remove constraints \& merge all positive interactions}
      \State Push signed interactions $\interact{}(S \cup S')$ to PQ, $\forall S, S' \in \Pi$
      \Repeat 
        \State Pop $S,S'\in\Pi$ with highest interaction $\interact{}(S \cup S')$ from PQ
        \If{\big[{\color{green}\textbf{$\interact{}(S \cup S') > 0$}}\big]}
          \State Merge clusters $S$, $S'$ and update hierarchical tree $T^*$
          \State Update interactions with neighboring clusters
        \EndIf
      \Until{\big[{\color{red}\textbf{$\interact{}(S \cup S') \leq 0$}}\big]} 
      \State Save the final clustering $\Pi^* \gets \Pi$ 
      \State \emph{// Phase 3: Merge negative interactions until one single cluster is left}
      \Repeat
        \State Pop $S,S'\in\Pi$ with highest interaction $\interact{}(S \cup S')$ from PQ 
        \State Merge clusters $S$, $S'$ and update hierarchical tree $T^*$
        \State Update interactions with neighboring clusters
      \Until{\big[Only one cluster is left in $\Pi$\big]} 
      % \EndIf
      \State
      \Return $\Pi^*$, $T^*$
  \end{algorithmic}
    \label{main_alg}
  \end{flushleft}

@@

\paragraph{Average, Single, and Complete Linkage} -- These three linkage criteria have been thoroughly studied on unsigned graphs, but never - until very recently - on signed graphs. In concurrent independent related work \cite{chehreghani2020hierarchical}, the authors prove that applying these three linkage methods to a signed graph is equivalent to applying them to the unsigned graph obtained by shifting all edge weights by a constant.  Here, we prove which of the algorithms studied here are ``intrinsically signed'' and do not have this invariance-property:
\begin{restatable}{prop}{invariantAlgs}
\label{prop:weight_shift_invariant}
We call an agglomerative algorithm ``weight-shift invariant'' if the dendrogram $T$ returned by the algorithm is invariant w.r.t. a shift of all edge weights $w_e$  by a constant $\alpha\in \mathbb{R}$. Among the variations of \algname{}, only hierarchical clustering with Average (HC-Avg), Single (HC-Single), and Complete linkage (HC-Complete) are weight-shift-invariant (see green box in Table~\ref{tab:linkage-criteria}).
\end{restatable}
\noindent Although average and single linkage methods have been largely studied on unsigned graphs, to our knowledge, they have never been combined with cannot-link constraints on signed graphs\footnote{Note that Complete linkage methods return the same clustering whether constraints are enforced or not (proof in Lemma \ref{lemma:absMax_and_complete_property}, in Appendix).}, so we name these algorithms \emph{HCC-Avg} and \emph{HCC-Single}. 
% On the other hand, we prove that an agglomerative algorithm with Complete linkage (HC-Complete) returns the same clustering whether cannot-link-constraints are enforced or not (Lemma \ref{lemma:absMax_and_complete_property}, in Appendix).
% , we prove that with Complete linkage the agglomeration order does not change whether constraints are introduced or not.
% \TODO{} in terms of scores, the advantage of one linkage over the other seems to be application dependent, but the median linkage has higher runtime.

% Minor objection: in section 3.1, the input graph is undirected, i.e., E \subseteq {V\choose 2}. Therefore, (S\times S')\cap E should be {{S\cup S'}\choose 2} \cap E. Note, this are edges of the subgraph induced by nodes S\cup S'. In fact, writing W(S,S') is confusing because W depends only on S\cup S', so I'd rather write W(S\cup S').

@@

\paragraph{Algorithms defining an ultrametric} -- 
GASP cannot be seen as a local search algorithm of the multicut optimization problem \ref{eq:MC_objective} (see more details in Appendix \ref{sec:relation_to_multicut}).
However, the connection between agglomerative algorithms and ultrametrics\footnote{A metric space $(X,d)$ is an \emph{ultrametric} if, for every $x,y,z \in X$, $d(x,y)\leq \max \{d(x,z), d(y,z)\}$.} is well known. Usually, ultrametrics are associated to strictly positive \emph{similarity} or \emph{dissimilarity} measures on a graph. In our framework, a trivial ultrametric is always given by the height $h_T$ of the dendrogram. However, for some of the \algname{} variations, we now define an ultrametric based on the edge weights and the signed interactions between clusters, generalizing what has been done for HC on unsigned graphs \cite{johnson1967hierarchical,milligan1979ultrametric}. To define this measure and prove its ultrametric property, we first map the signed interaction $\treeHeight{}$ defined in Eq.~\ref{eq:def_dendr_interact} to positive ``pseudo-distances'' $d_{T}:V \times V \rightarrow \mathbb{R}^{+}$:
\begin{align}\label{eq:UM_def}
d_{T}(u,v)\equiv &\begin{cases}
0 & \text{if}\,\, u=v\\
M-\treeHeight(u,v) & \text{if}\,\, u\neq v
\end{cases}\quad \forall u,v\in V\\
\mathrm{where}& \quad  M \equiv  \epsilon + \max_{u',v'\in V,\,u'\neq v'}\treeHeight(u',v')
\end{align}
and where $\epsilon > 0$. 
% By definition, we have that $d_{T}(u,v)=d_{T}(v,u)$ and $d_{T}(u,u)= 0$. 
We then prove the following proposition:
\begin{restatable}{prop}{secondUltraMetricProperty}
\label{prop:ultraMetric2}
% Consider the binary rooted tree $T^*$ returned by the \TODO{all variants in Table 1} \algname{} Algorithm~\ref{main_alg}, given a choice of linkage criteria $\interact{}$ and the use (or not) of CannotLinkConstraints. 
Among the algorithms included in the \algname{} framework (see Table \ref{tab:linkage-criteria}), only Mutex Watershed and hierarchical clustering with Average (HC-Avg), Single (HC-Single) and Complete linkage (HC-Complete) define an ultrametric $(V, d_{T^*})$, where $d_{T^*}$ is defined in Eq.~\ref{eq:UM_def} and $T^*$ is the tree returned by the \algname{} Algorithm~\ref{main_alg}.
\end{restatable}
\noindent In summary, in this section we have extended the family of HC algorithms \cite{johnson1967hierarchical,milligan1979ultrametric} with ``weight-based ultrametrics'' to signed graphs. Next, we move to their empirical evaluation.

### 2022-05-19 chapters/5_experiments_intro.tex
Update to CVPR2022 style
Camera ready changes

\section{Experiments}\label{sec:neuro_segm_exp}
\subsection{Signed graph clustering problems} \label{sec:clustering_problems}
We evaluate the agglomerative clustering algorithms included in our framework on a large collection of both synthetic and real-world graphs with very different structures. The size of the graphs ranges from a few hundred to hundreds of millions of edges.

@@

    \caption{\textbf{Tables (a-b)}: Scores and run times of algorithms in the \algname{} framework on the \emph{CREMI-gridGraph} and \emph{CREMI-3D-RAG} clustering problems: average linkage methods achieved the best accuracy. Measures shown are: Adapted-Rand error (ARAND, lower is better); Variation of Information (VOI) \cite{arganda2015crowdsourcing} (VOI-merge for under-clustering error and VOI-split for over-clustering error, lower values are better). \textbf{Table (c)}: Current leading entries in the CREMI challenge leaderboard (March 2021). CREMI-score is given by the geometric mean of (VOI-split + VOI-merge)  and ARAND error (lower is better).}

@@

\textbf{GASP on synthetic SSBM graphs}  -- 
% can be connected to any  that, depending on the edge-probability parameter $p$, are also \emph{dense}. 
\algname{} algorithms using cannotLinkConstraints are not expected to perform well on these graphs, because of the type of employed sign noise, so we focus our comparison only on the GAEC, HC-Avg and MWS algorithms (using Sum, Average, and AbsMax linkage methods, respectively).
Empirically, we observe that GAEC is the agglomerative algorithm performing best on SSBM graphs, on par with spectral method SPONGE$_{sym}$ (see Fig.~\ref{fig:SSBM_scores}). Given the simple properties of SSBM graphs, we can now give a detailed explanation of these empirical results. 
In SSBM graphs, the number of edges $E_{ij}\equiv(S_i\times S_j)\cap E$ connecting two clusters $S_i,S_j$ is  proportional to the product $|S_i|\cdot|S_j|$ of cluster sizes. 
With Sum or Avg linkage methods, due to the law of large numbers, the flipping noise is ``averaged out'' as soon as the set $E_{ij}$ becomes larger and clusters grow in size.
On the other hand, when clusters are small, it can happen that, for few clusters, several of their edges in $E_{ij}$ are flipped and the algorithm makes a mistake by merging two clusters belonging to different ground truth communities. From this observation, it follows that the \emph{flooding strategy} of the sum-linkage algorithm GAEC is a very good strategy on these types of graphs, because clusters are immediately grown in size (see dendrograms in Fig.~\ref{fig:dendrograms}). Average linkage method HC-Avg instead performs much worse on these graphs because it grows small equally-sized clusters and makes several wrong merge-decisions at the beginning. 
% This observation also explains the empirical results of \cite{chehreghani2020hierarchical}, which involved graph instances with very similar structure to the ones considered here. 
Lastly, the MWS algorithm is not expected to perform well on these graphs because of the high sensitivity of the AbsMax linkage to flipping noise. In Proposition \ref{prop:MWS_on_SSBM} (see Appendix), we prove that, at every iteration, the MWS algorithm makes a mistake with at least probability $\eta$, independently on the sizes of the two clusters that are popped from priority queue.
In summary, for the SSBM, we can obtain a deep understanding of the dynamics induced by various linkage criteria, and find that GAEC gives highest accuracy by a large margin.

@@

\begin{figure}
\centering
\includegraphics[width=0.48\textwidth,trim=0 10 60 0, clip]{./figs/comparison_GAEC_MWS_Avg.png} % left bottom right top
\caption{Failure cases of three versions of \algname{} applied to neuron segmentation. Only \emph{wrongly} segmented regions are highlighted in different warm colors. Red arrows point to wrongly split regions; yellow arrows point to false merge errors. HC-Avg returned the best segmentation. Data is 3D, hence the same color could be assigned to parts of segments that appear disconnected in 2D.  
\label{fig:failure_cases}}
\end{figure}

>>>>

\section{Experiments}\label{sec:neuro_segm_exp}
\subsection{Signed graph clustering problems} \label{sec:clustering_problems}
We evaluate the agglomerative clustering algorithms included in our framework on a large collection of both synthetic and real-world graphs with very different structures. The size of the graphs ranges from a few hundred to hundreds of millions of edges. In this way, we will highlight the strengths and limitations of the different linkage criteria introduced in the last section.

@@

    \caption{\textbf{Tables (a-b)}: Scores and run times of algorithms in the \algname{} framework on the \emph{CREMI-gridGraph} and \emph{CREMI-3D-RAG} clustering problems: average linkage methods achieved the best accuracy. Measures shown are: Adapted-Rand error (ARAND, lower is better); Variation of Information (VOI) \cite{arganda2015crowdsourcing} (VOI-merge for under-clustering error and VOI-split for over-clustering error, lower values are better). \textbf{Table (c)}: Current leading entries in the CREMI challenge leaderboard (November 2021). CREMI-score is given by the geometric mean of (VOI-split + VOI-merge)  and ARAND error (lower is better).}

@@

\textbf{GASP on synthetic SSBM graphs}  -- 
% can be connected to any  that, depending on the edge-probability parameter $p$, are also \emph{dense}. 
\algname{} algorithms using cannotLinkConstraints are not expected to perform well on these graphs, because of the type of employed sign noise, so we focus our comparison only on the GAEC, HC-Avg and MWS algorithms (using Sum, Average, and AbsMax linkage methods, respectively).
Empirically, we observe that GAEC is the agglomerative algorithm performing best on SSBM graphs, on par with spectral method SPONGE$_{sym}$ (see Fig.~\ref{fig:SSBM_scores}). Given the simple properties of SSBM graphs, we can now give a detailed explanation of these empirical results. 
In SSBM graphs, the number of edges $E_{ij}$
% \equiv(S_i\times S_j)\cap E$ 
connecting two clusters $S_i,S_j$ is  proportional to the product $|S_i|\cdot|S_j|$ of cluster sizes. 
With Sum or Avg linkage methods, due to the law of large numbers, the flipping noise is ``averaged out'' as soon as the set $E_{ij}$ becomes larger and clusters grow in size.
On the other hand, when clusters are small, it can happen that, for few clusters, several of their edges in $E_{ij}$ are flipped and the algorithm makes a mistake by merging two clusters belonging to different ground truth communities. From this observation, it follows that the \emph{flooding strategy} of the sum-linkage algorithm GAEC is a very good strategy on these types of graphs, because clusters are immediately grown in size (see dendrograms in Fig.~\ref{fig:dendrograms}). Average linkage method HC-Avg instead performs much worse on these graphs because it grows small equally-sized clusters and makes several wrong merge-decisions at the beginning. 
% This observation also explains the empirical results of \cite{chehreghani2020hierarchical}, which involved graph instances with very similar structure to the ones considered here. 
Lastly, the MWS algorithm is not expected to perform well on these graphs because of the high sensitivity of the AbsMax linkage to flipping noise. In Proposition \ref{prop:MWS_on_SSBM} (see Appendix), we prove that, at every iteration, the MWS algorithm makes a mistake with at least probability $\eta$, independently on the sizes of the two clusters that are popped from priority queue.
In summary, for the SSBM, we can obtain a deep understanding of the dynamics induced by various linkage criteria, and find that GAEC gives highest accuracy by a large margin.

@@

\begin{figure}
\centering
\includegraphics[width=0.48\textwidth,trim=0 10 60 0, clip]{./figs/comparison_GAEC_MWS_Avg.pdf} % left bottom right top
\caption{Failure cases of three versions of \algname{} applied to neuron segmentation. Only \emph{wrongly} segmented regions are highlighted in different warm colors. Red arrows point to wrongly split regions; yellow arrows point to false merge errors. HC-Avg returned the best segmentation. Data is 3D, hence the same color could be assigned to parts of segments that appear disconnected in 2D.  
\label{fig:failure_cases}}
\end{figure}
